Notes on conformal manifolds

### 2025-05-16 Draft_MLFlatDirections.tex
Little comments.
Comments during meeting.

As a first simplified example we have considered only the parameters $x_{1}, x_{2}, x_{4}, x_{8}$ and $x_{10}$. The potential~\eqref{eq:scalarpotential} becomes
\begin{equation} \label{eq:scalarpotential124810}
	\begin{aligned}
		V &= \frac{1}{8}\, e^{-2\,x_{8}} \bigg[4 + 4\,x_{1}^4 + e^{4\,x_{8}} \big(2 + x_{10}^2\big)^2 \big(1 + x_{2}^4\big)\\
		  & \quad\qquad\qquad - 4\,e^{3\,x_{8}}\,\big(2 + x_{10}^2\big) \Big(2 - x_{1}^2 + \sqrt{2}\,x_{1} x_{10} x_{2}^3 + x_{2}^4 + 
      			2\,x_{1} x_{2} x_{4} + x_{4}^2 - x_{2}^2 \big(1 - x_{1}^2 + x_{4}^2\big)\Big)\\
        & \quad\qquad\qquad - 8\,e^{x_{8}}\,\Big(2 + x_{1}^4 + \sqrt{2}\,x_{1}^3 x_{10} x_{2} - x_{2}^2 - 2\,x_{1} x_{2} x_{4} + 
      			x_{4}^2 - x_{1}^2 \big(1 - x_{2}^2 + x_{4}^2\big)\Big)\\
        & \quad\qquad\qquad + 4\,e^{2\,x_{8}}\,\Big(2 - 4\,x_{1}^2 + x_{1}^4 - 4\,x_{2}^2 + 4\,x_{1}^2 x_{2}^2 + x_{2}^4
        			+ x_{10}^2\big(1 + 3\,x_{1}^2 x_{2}^2\big) + 4\,x_{4}^2 + x_{4}^4\Big)\\
        & \quad\qquad\qquad + 8\sqrt{2}\,x_{10}\,e^{2\,x_{8}}\,\Big(x_{1}^3 x_{2} + x_{1}^2 x_{4} - x_{2}^2 x_{4} + x_{1} \big(x_{2}^3 - x_{2} x_{4}^2\big)\Big)\bigg].
	\end{aligned}
\end{equation}

\section{Numerical analysis}
	\subsection{Gradient descent and local analysis}
	\begin{itemize}[label=\textbullet]
		\item Gradient descent: implementation, sampling, time needed, loss.

		In order to look for flat direction in this potential, we will use numerical tools. The path we will follow is the following : first we will sample the underlying manifold, then we will use numerical tools to exctract analytics information on this cloud of points. Therefore the first step is to sample the manifold. To do so, we choose to do a basic gradient descent. We start by randomly uniformy initializing points in an hypercube of length $[-2,2]$. This will be important latter to have points which are not only in $[-1,1]$. We choose to have $10^5$ points. For this example. This number is motivated by the dimensionality of our problem. If all possible directions are going to be populated, and if wen want $\mathcal{O}(10)$ points in each direction, then we need  $\mathcal{O}(10^5)$ points. As the true internal dimension of the manifold is lesser or equal than 5, this number provides an upper bound for the number of points needed to sufficiently sample the manifold.  We then perfom a gradient descent on the point using the automatic differentiation method of tensorflow. For the loss function, we use 
		\begin{equation}
			\mathcal{L} = \sum_{i = 1}^{n_{points}} |\nabla V(\vec{x}_i)|^2
		\end{equation}
		with $\vec{x} = (x_1,x_2,x_4,x_8,x_{10})$. We choose to use the Adam optimizer and use a learning rate of $10^{-2}$. We run the gradient descent for $10000$ epochs to make sure of convergence. 
%		\begin{figure}
%			\includegraphics[scale = 0.5]{imagefile} % add loss history 
%		\end{figure}
		As a first visualisation, we show a triangular plot of the data in Fig. \ref{triangular_plot_1_2_4_8_10}. This shows all possible $2d$ projections of the data as well as the $1d$ histograms of the data after the gradient descent. 
		\begin{figure}
			\centering
			\includegraphics[scale = 0.25]{Figures/triangular_plot_1_2_4_8_10.png} %names need to be changed.
			\caption{Triangular plot showing $2d$ projections and $1d$ histograms of the data after the gradient descent}
			\label{triangular_plot_1_2_4_8_10}
		\end{figure}

		Comments :
		\begin{enumerate}
			\item We see some non trivial correlation between the data. Note however that some features we see can be caused by the fact that some places may be greater basin of attractions (this is for example what happend in the $x_1/x_2$ plot). However, the structure we see in the $x_4/x_8$ or $x_4/x_{10}$ is a remnant of true features of the manifold
			\item All directions seems to be well populated 
			\item A great majority of the points fall in the $\vec{0}$ point. This is not too surprising. This just means that the round sphere solution has a greater basin of attraction. Even so, those points are going to be sufficiently good for the rest of the analysis, and we do not try to move the points along flat directions (to depopulate the origin et repopulate the rest).
		\end{enumerate}
		\item Local PCA: graphs of number of patches with given local dimension, and with varying size of local patch, discuss optimal patch size (we want at least fex points in each direction).

@@

		\item Clusteting: HDBScan (only parameter: minimal number of points in a cluster), and graphs of 3d slices of the 5d space.

		Something that might happen, is that we have several manifolds of dimesion 3, and the dimension 4 points we saw above are actually the intersection of those manifolds. If we indeed picture the intersection of two lines, then at the intersection point, the local dimension found by the previous algorithm should be 2. To exclude such a scenario, we run a clustering algorithm, based on local density. For this, we use the HDBSAN algorithm. (details on how it works?). We choose a minimal cluster size of 10 points. Out of the 100000, the algorithm finds 3 cluster : one on dimension 10, one of dimension 91747 and a last one of dimension 6849 (not a total of 100000??). The last cluster are just the points that the algorithm did not managed to fit into any cluster. For the first one, we interpret it as some fluctuation in our data, making some region of different density as the other, but we do not believe this is some real result, more an artefact. Therefore, what the algorithm tells us, is that there seems to be on cluster with more than $90\%$ of the data. We show on \ref{3dplots} three-dimensional projections of the data.

@@

		%\item After PCA and clustering we know that the points after the gradient descent sample a three-dimensional manifold.
	\end{itemize}

@@

	\paragraph{Results}
	We have used the numerical analysis outlined in the previous sections to the search for extrema of the scalar potential~\eqref{eq:scalarpotential124810}. For 100 runs with parameters
	\begin{equation}
		\begin{gathered}
			n_{\rm iter} = 1000, \quad n_{\rm particles} = 1000, \quad {\rm reg} = 10^{3}, \quad \beta = 10^{-10} + \left(\frac{i}{n_{\rm iter}}\right)^{5}, \quad \sigma = 0.5 - \frac{0.45}{1+\exp\big(10-20 i/n_{\rm iter}\big)}, \\
			{\rm sample\ size = 10000},\quad p_{\rm add} = 0, \quad p_{\rm remove} = 0, \quad p_{\rm modify} = 0.5, \quad p_{\rm multiply} = 0.25, \quad p_{\rm divide} = 0.25,
		\end{gathered}
	\end{equation}
	the code finds the 7 different following polynomials:
	\begin{subequations} \label{eq:pols}
    \begin{align}
      p_{1} &= -\sqrt{2}\,x_{1} + \sqrt{2}\,x_{1}\tilde{x}_{8} + x_{2}\tilde{x}_{8}x_{10} - \sqrt{2}\,x_{2}x_{4}\tilde{x}_{8},\\
      p_{2} &= 2\,x_{2} + 2\,x_{2}\tilde{x}_{8} + \sqrt{2}\,x_{1}x_{10} + 2\,x_{1}x_{4} - x_{2}\tilde{x}_{8}x_{10}^{2},\\
      p_{3} &= 2\,x_{2} - 2\,x_{2}\tilde{x}_{8} + \sqrt{2}\,x_{1}\tilde{x}_{8}x_{10} + 2\,x_{1}x_{4}\tilde{x}_{8}
      		- 2\,x_{2} x_{4}^{2}\tilde{x}_{8},\\
      p_{4} &= \sqrt{2}\,x_{2} - \sqrt{2}\,x_{2}\tilde{x}_{8} + \sqrt{2}\,x_{1}\,x_{4} + x_{1}\tilde{x}_{8}x_{10}
      		- x_{2}x_{4}\tilde{x}_{8}x_{10},\\
      p_{5} &= -2\,x_{1} - 2\,x_{2}x_{4}- 2\,x_{1}x_{4}^{2} + 2\,x_{1}\tilde{x}_{8} + \sqrt{2}\,x_{2}x_{10}
      		+ x_{1}\tilde{x}_{8}x_{10}^{2},\\
      p_{6} &= -2\,x_{2}x_{4} - 2\,x_{1}x_{4}^{2} + 2\,x_{2}x_{4}\tilde{x}_{8} + \sqrt{2}\,x_{2}x_{10} 
      		- \sqrt{2}\,x_{2}\tilde{x}_{8}x_{10} +  x_{1}\tilde{x}_{8}x_{10}^{2},\\
      p_{7} &= -\sqrt{2}\,x_{1}^{2}x_{4} + \sqrt{2}\,x_{2}^{2}x_{4} + \sqrt{2}\,x_{1}x_{2}x_{4}^{2} - x_{1}^{2}x_{10}
      		- x_{2}^{2}x_{10},
    \end{align}
  \end{subequations}
  where $\tilde{x}_{8} = e^{x_{8}}$, with the distribution:
  \begin{equation}
  	\begin{tabular}{l|cccccccc}
  		& $p_{1}$ & $p_{2}$ & $p_{3}$ & $p_{4}$ & $p_{5}$ & $p_{6}$ & $p_{7}$ & $\varnothing$ \\\hline\hline
  		Distribution & 59\% & 25\% & 5\% & 9\% & 2\% & 4\% & 2\% & 13\%
  	\end{tabular}
  \end{equation}
  The sum of the percentages is higher than 100\% because some runs find more than one polynomials. The code fails to find any polynomial in 13\% of the time. The averaged time needed for a single run on a PC in $\sim 1400\,{\rm s}$.

@@

For every couples $(p_{i},p_{j})$ of polynomials in eq.~\eqref{eq:pols}, the system
\begin{equation}
	\begin{cases}
		p_{i} = 0, \\
		p_{j} = 0,
	\end{cases}
\end{equation}
gives the same solution
\begin{equation}
	\begin{cases}
		\displaystyle x_{1} = \frac{x_{2}}{\sqrt{2}}\,\frac{e^{x_{8}/2}}{e^{x_{8}}-1}\,\Big(-x_{5}\,e^{x_{8}/2} + \sqrt{2-4\,e^{x_{8}}+e^{2x_{8}}\big(2+x_{10}^{2}\big)}\Big),\\[8pt]
		\displaystyle x_{4} = \frac{e^{-x_{8}/2}}{\sqrt{2}}\,\sqrt{2-4\,e^{x_{8}}+e^{2x_{8}}\big(2+x_{10}^{2}\big)}.
	\end{cases}
\end{equation}
We checked that this defines a 3-parameter solution of three-dimensional half-maximal supergravity. Equivalently:
\begin{equation}
	\begin{cases}
		\displaystyle \tilde{x}_{8} = \frac{x_{1}^{2}+x_{2}^{2}}{x_{2}^{2} + \big(x_{1}-x_{2}x_{4}\big)^{2}},\\[10pt]
		\displaystyle x_{10} = \sqrt{2}\,x_{4}\,\frac{x_{2}^{2} - x_{1}^{2}+x_{1}x_{2}x_{4}}{x_{1}^{2}+x_{2}^{2}}.
	\end{cases}
\end{equation}
The solution preserves a ${\rm U}(1)\times{\rm U}(1)$ gauged symmetry.

>>>>

As a first simplified example we have considered only the parameters $x_{1}, x_{2}, x_{4}, x_{8}$ and $x_{10}$. The potential~\eqref{eq:scalarpotential} becomes
\begin{equation} \label{eq:scalarpotential124810}
	\begin{aligned}
		V &= \frac{1}{8}\, e^{-2\,x_{8}} \bigg[4 + 4\,x_{1}^4 + e^{4\,x_{8}} \big(2 + x_{10}^2\big)^2 \big(1 + x_{2}^4\big)\\
		  & \quad\qquad\qquad - 4\,e^{3\,x_{8}}\,\big(2 + x_{10}^2\big) \Big(2 - x_{1}^2 + \sqrt{2}\,x_{1} x_{10} x_{2}^3 + x_{2}^4 + 
      			2\,x_{1} x_{2} x_{4} + x_{4}^2 - x_{2}^2 \big(1 - x_{1}^2 + x_{4}^2\big)\Big)\\
        & \quad\qquad\qquad - 8\,e^{x_{8}}\,\Big(2 + x_{1}^4 + \sqrt{2}\,x_{1}^3 x_{10} x_{2} - x_{2}^2 - 2\,x_{1} x_{2} x_{4} + 
      			x_{4}^2 - x_{1}^2 \big(1 - x_{2}^2 + x_{4}^2\big)\Big)\\
        & \quad\qquad\qquad + 4\,e^{2\,x_{8}}\,\Big(2 - 4\,x_{1}^2 + x_{1}^4 - 4\,x_{2}^2 + 4\,x_{1}^2 x_{2}^2 + x_{2}^4
        			+ x_{10}^2\big(1 + 3\,x_{1}^2 x_{2}^2\big) + 4\,x_{4}^2 + x_{4}^4\Big)\\
        & \quad\qquad\qquad + 8\sqrt{2}\,x_{10}\,e^{2\,x_{8}}\,\Big(x_{1}^3 x_{2} + x_{1}^2 x_{4} - x_{2}^2 x_{4} + x_{1} \big(x_{2}^3 - x_{2} x_{4}^2\big)\Big)\bigg].
	\end{aligned}
\end{equation}
\ce{Change notation to keep $x$ for the polynomial variables?}

\ce{Discuss here why we need numerical methods: to complicated for mathematica to simplify the gradient.}

\section{Numerical analysis}
	\subsection{Gradient descent and local analysis}
		\paragraph{Gradient descent: implementation, sampling, time needed, loss}
		In order to look for flat direction in this potential, we will use numerical tools. The path we will follow is the following : first we will sample the underlying manifold, then we will use numerical tools to exctract analytics information on this cloud of points. Therefore the first step is to sample the manifold. To do so, we choose to do a basic gradient descent. We start by randomly uniformy initializing points in an hypercube of length $[-2,2]$. This will be important latter to have points which are not only in $[-1,1]$. We choose to have $10^5$ points. For this example. This number is motivated by the dimensionality of our problem. If all possible directions are going to be populated, and if wen want $\mathcal{O}(10)$ points in each direction, then we need  $\mathcal{O}(10^5)$ points. As the true internal dimension of the manifold is lesser or equal than 5, this number provides an upper bound for the number of points needed to sufficiently sample the manifold.  We then perfom a gradient descent on the point using the automatic differentiation method of tensorflow. For the loss function, we use 
		\begin{equation}
			\mathcal{L} = \sum_{i = 1}^{n_{points}} |\nabla V(\vec{x}_i)|^2
		\end{equation}
		with $\vec{x} = (x_1,x_2,x_4,x_8,x_{10})$. We choose to use the Adam optimizer and use a learning rate of $10^{-2}$. We run the gradient descent for $10000$ epochs to make sure of convergence. 
%		\begin{figure}
%			\includegraphics[scale = 0.5]{imagefile} % add loss history 
%		\end{figure}
		As a first visualisation, we show a triangular plot of the data in Fig. \ref{triangular_plot_1_2_4_8_10}. This shows all possible $2d$ projections of the data as well as the $1d$ histograms of the data after the gradient descent. 
		\begin{figure}[t]
			\centering
			\includegraphics[scale = 0.25]{Figures/triangular_plot_1_2_4_8_10.png} %names need to be changed.
			\caption{Triangular plot showing $2d$ projections and $1d$ histograms of the data after the gradient descent}
			\label{triangular_plot_1_2_4_8_10}
		\end{figure}

		Comments :
		\begin{enumerate}
			\item We see some non trivial correlation between the data. Note however that some features we see can be caused by the fact that some places may be greater basin of attractions (this is for example what happend in the $x_1/x_2$ plot). However, the structure we see in the $x_4/x_8$ or $x_4/x_{10}$ is a remnant of true features of the manifold
			\item All directions seems to be well populated 
			\item A great majority of the points fall in the $\vec{0}$ point. This is not too surprising. This just means that the round sphere solution has a greater basin of attraction. Even so, those points are going to be sufficiently good for the rest of the analysis, and we do not try to move the points along flat directions (to depopulate the origin et repopulate the rest).
		\end{enumerate}

		\paragraph{Local PCA: graphs of number of patches with given local dimension, and with varying size of local patch, discuss optimal patch size (we want at least fex points in each direction)}

@@

		\paragraph{Clusteting: HDBScan (only parameter: minimal number of points in a cluster), and graphs of 3d slices of the 5d space}
		Something that might happen, is that we have several manifolds of dimesion 3, and the dimension 4 points we saw above are actually the intersection of those manifolds. If we indeed picture the intersection of two lines, then at the intersection point, the local dimension found by the previous algorithm should be 2. To exclude such a scenario, we run a clustering algorithm, based on local density. For this, we use the HDBSAN algorithm. (details on how it works?). We choose a minimal cluster size of 10 points. Out of the 100000, the algorithm finds 3 cluster : one on dimension 10, one of dimension 91747 and a last one of dimension 6849 (not a total of 100000?? Clustering only on the point with 3d neighbourhood to avoid intersections). The last cluster are just the points that the algorithm did not managed to fit into any cluster. For the first one, we interpret it as some fluctuation in our data, making some region of different density as the other, but we do not believe this is some real result, more an artefact. Therefore, what the algorithm tells us, is that there seems to be on cluster with more than $90\%$ of the data. We show on \ref{3dplots} three-dimensional projections of the data.

@@

		After PCA and clustering we know that the points after the gradient descent sample a three-dimensional manifold.

@@

	\paragraph{Results}
	We have used the numerical analysis outlined in the previous sections to the search for extrema of the scalar potential~\eqref{eq:scalarpotential124810}. For 100 runs with parameters \ce{\texttt{max degree} is 4, \texttt{max num monomials} is 6.}
	\begin{equation}
		\begin{gathered}
			n_{\rm iter} = 1000, \quad n_{\rm particles} = 1000, \quad {\rm reg} = 10^{3}, \quad \beta = 10^{-10} + \left(\frac{i}{n_{\rm iter}}\right)^{5}, \quad \sigma = 0.5 - \frac{0.45}{1+\exp\big(10-20 i/n_{\rm iter}\big)}, \\
			{\rm sample\ size = 10000},\quad p_{\rm add} = 0, \quad p_{\rm remove} = 0, \quad p_{\rm modify} = 0.5, \quad p_{\rm multiply} = 0.25, \quad p_{\rm divide} = 0.25,
		\end{gathered}
	\end{equation}
	the code finds the 7 different following polynomials: \ce{Mention that we ``rounded'' the coefficients to integer and square root of them.}
	\begin{subequations} \label{eq:pols}
    \begin{align}
      p_{1} &= -\sqrt{2}\,x_{1} + \sqrt{2}\,x_{1}\tilde{x}_{8} + x_{2}\tilde{x}_{8}x_{10} - \sqrt{2}\,x_{2}x_{4}\tilde{x}_{8},\\
      p_{2} &= 2\,x_{2} + 2\,x_{2}\tilde{x}_{8} + \sqrt{2}\,x_{1}x_{10} + 2\,x_{1}x_{4} - x_{2}\tilde{x}_{8}x_{10}^{2},\\
      p_{3} &= 2\,x_{2} - 2\,x_{2}\tilde{x}_{8} + \sqrt{2}\,x_{1}\tilde{x}_{8}x_{10} + 2\,x_{1}x_{4}\tilde{x}_{8}
      		- 2\,x_{2} x_{4}^{2}\tilde{x}_{8},\\
      p_{4} &= \sqrt{2}\,x_{2} - \sqrt{2}\,x_{2}\tilde{x}_{8} + \sqrt{2}\,x_{1}\,x_{4} + x_{1}\tilde{x}_{8}x_{10}
      		- x_{2}x_{4}\tilde{x}_{8}x_{10},\\
      p_{5} &= -2\,x_{1} - 2\,x_{2}x_{4}- 2\,x_{1}x_{4}^{2} + 2\,x_{1}\tilde{x}_{8} + \sqrt{2}\,x_{2}x_{10}
      		+ x_{1}\tilde{x}_{8}x_{10}^{2},\\
      p_{6} &= -2\,x_{2}x_{4} - 2\,x_{1}x_{4}^{2} + 2\,x_{2}x_{4}\tilde{x}_{8} + \sqrt{2}\,x_{2}x_{10} 
      		- \sqrt{2}\,x_{2}\tilde{x}_{8}x_{10} +  x_{1}\tilde{x}_{8}x_{10}^{2},\\
      p_{7} &= -\sqrt{2}\,x_{1}^{2}x_{4} + \sqrt{2}\,x_{2}^{2}x_{4} + \sqrt{2}\,x_{1}x_{2}x_{4}^{2} - x_{1}^{2}x_{10}
      		- x_{2}^{2}x_{10},
    \end{align}
  \end{subequations}
  where $\tilde{x}_{8} = e^{x_{8}}$, with the distribution: \ce{It's not a distribution, how to name it? Frequency?}
  \begin{equation}
  	\begin{tabular}{l|cccccccc}
  		& $p_{1}$ & $p_{2}$ & $p_{3}$ & $p_{4}$ & $p_{5}$ & $p_{6}$ & $p_{7}$ & $\varnothing$ \\\hline\hline
  		Distribution & 59\% & 25\% & 5\% & 9\% & 2\% & 4\% & 2\% & 13\%
  	\end{tabular}
  \end{equation}
  The sum of the percentages is higher than 100\% because some runs find more than one polynomials. The code fails to find any polynomial in 13\% of the time. The averaged time needed for a single run on a PC in $\sim 1400\,{\rm s}$. \ce{This timing is with \texttt{analysis}, it may be less without.}

  \ce{Draw histograms of the polynomials distribution to discuss runs with multiple outputs?}

@@

For every couples $(p_{i},p_{j})$ of polynomials in eq.~\eqref{eq:pols}, the system
\begin{equation}
	\begin{cases}
		p_{i} = 0, \\
		p_{j} = 0,
	\end{cases}
\end{equation}
gives the same solution \ce{That's not true! For example, $x_{1}=x_{2}=0$ solves all polynomials but is not a sugra solution. Mention tests on the sugra?}
\begin{equation}
	\begin{cases}
		\displaystyle x_{1} = \frac{x_{2}}{\sqrt{2}}\,\frac{e^{x_{8}/2}}{e^{x_{8}}-1}\,\Big(-x_{5}\,e^{x_{8}/2} + \sqrt{2-4\,e^{x_{8}}+e^{2x_{8}}\big(2+x_{10}^{2}\big)}\Big),\\[8pt]
		\displaystyle x_{4} = \frac{e^{-x_{8}/2}}{\sqrt{2}}\,\sqrt{2-4\,e^{x_{8}}+e^{2x_{8}}\big(2+x_{10}^{2}\big)}.
	\end{cases}
\end{equation}
We checked that this defines a 3-parameter solution of three-dimensional half-maximal supergravity. Equivalently:
\begin{equation}
	\begin{cases}
		\displaystyle \tilde{x}_{8} = \frac{x_{1}^{2}+x_{2}^{2}}{x_{2}^{2} + \big(x_{1}-x_{2}x_{4}\big)^{2}},\\[10pt]
		\displaystyle x_{10} = \sqrt{2}\,x_{4}\,\frac{x_{2}^{2} - x_{1}^{2}+x_{1}x_{2}x_{4}}{x_{1}^{2}+x_{2}^{2}}.
	\end{cases}
\end{equation}
The solution preserves a ${\rm U}(1)\times{\rm U}(1)$ gauged symmetry.
